---
title: "p8130_hw5_xy2395"
author: "Jack Yan"
date: "11/30/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(faraway) 
library(patchwork)
library(funModeling)
```

# Dataset Description

`state.x77` is a matrix with 50 rows and 8 columns giving the following statistics in the respective columns:

*  Population: population estimate as of July 1, 1975.
*  Income: per capita income (1974).
*  Illiteracy: illiteracy (1970, percent of population).
*  Life Exp: life expectancy in years (1969–71).
*  Murder: murder and non-negligent manslaughter rate per 100,000 population (1976).
*  HS Grad: percent high-school graduates (1970).
*  Frost: mean number of days with minimum temperature below freezing (1931–1960) in capital or large city.
*  Area: land area in square miles.

```{r}
# Load the data
state_df = 
  state.x77 %>% 
  as.tibble() %>% 
  janitor::clean_names() %>% 
  select(life_exp, everything())
  
# state_df %>% skimr::skim()
```

# Problem 1: Exploratory Data Analysis

## Descriptive Statistics

Descriptive Statistics for all the variables are shown below. All the variables are continuous.

```{r}
# Build a function to generate descriptive statistics for continuous variables
summary_continuous = function(variable){
  data_frame(
    mean = mean(variable),
    sd = sd(variable),
    median = median(variable),
    maximum = max(variable),
    minimum = min(variable),
    IQR = IQR(variable)
  )
}

# Generate descriptive statistics
map(state_df, summary_continuous) %>% 
  bind_rows() %>% 
  mutate(variable = names(state_df)) %>% 
  select(variable, everything()) %>% 
  knitr::kable(digits = 2, 
               caption = "Descriptive statistics of continuous variables")
```

Also show the correlation matrix to check potential correlations. `life_exp` and `murder` are highly correlated. `illiteracy` and `murder` are correlated as well.

```{r}
cor(state_df) %>% 
   knitr::kable(digits = 2)
```

## Plots

First, plot the highly correlated variables.

```{r}
plot(state_df)

murder_illiteracy_points = 
state_df %>% 
  ggplot(aes(x = murder, y = illiteracy))+
    geom_point()
murder_lifeexp_points = 
state_df %>% 
  ggplot(aes(x = murder, y = life_exp))+
    geom_point()
murder_illiteracy_points + murder_lifeexp_points
```

Then check the normality of the outcome, `life_exp`. Life expectancies are approximately normally distributed among the states. 

```{r}
state_df %>% 
  ggplot(aes(x = life_exp)) +
  geom_histogram(bins = 15)
```

Also show the distribution of other variables.

```{r}
state_df %>% select(-life_exp) %>% 
plot_num()
```


# Problem 2: Model Building Using Automatic Procedures

